---
title: "First look at some tetraploids"
author: "David Gerard"
date: "`r Sys.Date()`"
---

## Abstract

Here, I look at some autotetraploids.

## Read in data

```{r}
ref <- read.table("../data/reference_david.txt")
alt <- read.table("../data/alternative_david.txt")
```

## Make a few plots

```{r, cache=TRUE}
library(updog)
for (index in 1:5) {
  ocounts <- ref[, index]
  osize   <- alt[, index] + ref[, index]
  print(plot_geno(ocounts, osize, 4))
}
```

```{r, cache=TRUE}
ulist <- list()
for (index in 1:5) {
  ocounts <- ref[, index]
  osize   <- alt[, index] + ref[, index]
  uout <- updog(ocounts = ocounts, osize = osize, model = "f1", ploidy = 4, print_val = FALSE)
  ulist[[index]] <- uout
  print(plot(uout))
}
```

It doesn't seem to fit too well. But the OD is huge in all of them, indicating that these data aren't too trustworthy.

```{r}
sapply(ulist, function(x) { x$od_param })
```

## Some thoughts

It seems that a lot of these SNP's have quite a lot of points really close to all AAAA and then everything else is dispersed. Indeed, almost every SNP has at least half of their reads being 95\% or more AAAA.

```{r}
library(tidyverse)
prop_mat <- ref / (ref + alt)
hist(colMeans(prop_mat > 0.95, na.rm = TRUE))
```

```{r info}
sessionInfo()
```
