---
title: "Diagnostic plot"
author: "David Gerard"
date: "May 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here I explore diagnostic plots


# Data

This is the problem SNP I have for updog.

```{r}
library(updog)
set.seed(483)
load("../data/subset_David.Rdata")
ddat <- subset_david
lastp <- 12 ## number of final rows that are parents

cmat <- ddat[[7]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]

p1counts <- p1dat[, 1]
p1size   <- rowSums(p1dat)
p2counts <- p2dat[, 1]
p2size   <- rowSums(p2dat)
ocounts  <- odat[, 1]
osize    <- rowSums(odat)
ploidy <- 6
```

Raw plots

```{r}
plot_geno(ocounts = ocounts, osize = osize, ploidy = ploidy, seq_error = 0)
```

# Fit updog

```{r, cache = TRUE}
uout <- updog_vanilla(ocounts = ocounts, osize = osize, ploidy = ploidy, print_val = FALSE,
                      update_outmean = FALSE)
```

```{r}
plot(uout, plot_beta = FALSE)
```

# EDA

My idea is to bin the genotypes and then see their discrepancy from the expecteed number under the segegation probabilities.

```{r}
tab_geno <- table(uout$ogeno)
phat <- tab_geno / sum(tab_geno)

qarray <- get_q_array(ploidy)
probs <- qarray[uout$p1geno + 1, uout$p2geno + 1, ]
```

Notice that `phat` deviates from 0.5 by the same amount. This is also the predicted segregation probabilities. Thought I checked other fits and they don't follow this. So I think this is just a coincidence.

```{r}
phat
probs[6:7]
```

```{r}
sessionInfo()
```
