---
title: "Initial Updog Fits"
author: "David Gerard"
date: "April 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I run the bias-modeling updog on the alfalfa data.

# Subset Data

```{r}
obj <- readRDS(file = "../output/allele_counts/chr1_ct.RDS")
str(obj)
```

I will choose SNP's that are at various levels of total read-counts.

```{r}
suppressMessages(library(tidyverse))
med_counts <- apply(obj$A + obj$a, 1, median, na.rm = TRUE)
mean_counts <- rowMeans(obj$A + obj$a, na.rm = TRUE)
quants <- quantile(mean_counts, probs = seq(0, 1, length = 30), na.rm = TRUE)
match_vec <- match(round(quants, digits = 1), round(mean_counts, digits = 1))
sub_A <- obj$A[match_vec, ]
sub_a <- obj$a[match_vec, ]
```

For the smaller counts alleles, most of the samples don't have any read-counts (indicated by `NA`) (though, this isn't a guarantee).

```{r}
plot(rowMeans(is.na(sub_a)), type = "l")
plot(rowMeans(is.na(sub_A)), type = "l")
```

# Raw Plots

First, some plots of the maximum expressed variants

```{r}
library(updog)
ploidy  <- 4
ocounts <- sub_A
osize   <- sub_A + sub_a

for (index in 1:nrow(ocounts)) {
 pl <- plot_geno(ocounts = ocounts[index, ], osize = osize[index, ],
                 ploidy = 4, seq_error = 0)  
 print(pl)
}

```

This last SNP makes no sense. Why are they all clustered at 0.75? Is this a technical artifact? This would only make sense biologically if segregation happened in pairs. E.g. if the parental genotypes were AAAA and ABAB and AB have to cluster together in the second parent. If they all segregated independently, then you would expect to see some AABB's and some AAAA's. This is weird. 

An alternative explanation is an extreme sequencing error rate combined with an extreme mapping bias toward the non-reference allele. This is what `updog_vanilla` catches.
```{r, cache = TRUE}
uout <- updog_vanilla(ocounts = ocounts[nrow(ocounts), ], osize = osize[nrow(osize), ],
                      ploidy = 4)
plot(uout, ask = FALSE)
```

Some of these don't look too bad!

```{r, cache = TRUE}
for (index in 20:29) {
  ocounts_nona <- ocounts[index, ]
  ocounts_nona <- ocounts_nona[!is.na(ocounts_nona)]
  osize_nona <- osize[index, ]
  osize_nona <- osize_nona[!is.na(osize_nona)]

  uout <- updog_vanilla(ocounts = ocounts_nona, osize = osize_nona, ploidy = ploidy, print_val = FALSE,
                        update_outdisp = FALSE, update_outmean = FALSE, update_outprop = TRUE,
                        non_mono_max = 1)
  plot(uout, ask = FALSE, plot_beta = FALSE)
  print(plot_geno(ocounts = ocounts_nona, osize = osize_nona, ploidy = 4, seq_error = 0))
}

```

```{r}
sessionInfo()
```
