---
title: "First SNP Analaysis"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This is a template for writing reports with R Markdown.

## Read in SNPs

```{r}
library(updog)
set.seed(3452)
load("../data/subset_David.Rdata")
dat <- subset_david
lastp <- 12 ## number of final rows that are parents
```

## Run `updog` iterative procedure on 10th SNP

We now run updog both with and without parental genotypes.
```{r}
cmat <- dat[[8]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]
uout1 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
               p1counts = p1dat[, 1], p1size = rowSums(p1dat),
               p2counts = p2dat[, 1], p2size = rowSums(p2dat), iterate = TRUE)

uout2 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
               iterate = TRUE)
uout3 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6)


```

The parental genotypes are chosen with near certainty
```{r}
round(uout1$p1postprob, digits = 2)
round(uout2$p1postprob, digits = 2)
round(uout3$p1postprob, digits = 2)

round(uout1$p2postprob, digits = 2)
round(uout2$p2postprob, digits = 2)
round(uout3$p2postprob, digits = 2)
```

Plot output
```{r}
## one at a time procedure
labelvec <- apply(uout1$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

## iterative procedure
labelvec <- apply(uout2$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

## uniform prior
labelvec <- apply(uout3$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)
```

## Uniform
What does a uniform prior on the parents mean for the prior on the children?

``` {r}
pvec <- rep(1/7, length = 7) ## prior on parents
qarray <- get_q_array(6)
harray <- sweep(qarray, MARGIN = 1, STATS = pvec, FUN = `*`)
harray <- sweep(harray, MARGIN = 2, STATS = pvec, FUN = `*`)

max(abs(qarray / 49 - harray))

hl <- apply(harray, 3, sum)
hl
```

So a uniform prior on the q-array is actually a pretty strong prior on the children.

Try a uniform prior on the children
```{r}
pout <- mapply(FUN = bin_post, odat[, 1], rowSums(odat), MoreArgs = list(prior = pvec))
labelvec <- apply(pout, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

pout <- mapply(FUN = bin_post, odat[, 1], rowSums(odat), MoreArgs = list(prior = hl))
labelvec <- apply(pout, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)
```

But the prior doesn't seem to change things that much.


## Session information

```{r info}
sessionInfo()
```
