---
title: "Local Minima"
author: "David Gerard"
date: "June 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Over-dispersion saves us because fewer local minima.

# Local Minima

The following is a simulated dataset that I had trouble on.

```{r}
library(updog)
rout <- readRDS("../data/rout.RDS")
plot(rout, plot_beta = FALSE)
```

We get a much smaller bias than what we should if we start at the defaults in `updog`.

```{r}
uout1 <- updog::updog_vanilla(ocounts = rout$input$ocounts, 
                              osize = rout$input$osize, 
                              ploidy = rout$input$ploidy, 
                              model = "hw",
                              out_prop = 0, 
                              update_outprop = FALSE)
plot(uout1, plot_beta = FALSE)
```

But if we start at the correct bias, we get good results

```{r}
uout2 <- updog::updog_vanilla(ocounts = rout$input$ocounts, 
                              osize = rout$input$osize, 
                              ploidy = rout$input$ploidy, 
                              model = "hw",
                              out_prop = 0, 
                              bias_val = rout$bias_val,
                              update_outprop = FALSE)
plot(uout2, plot_beta = FALSE)
```

The likelihood is higher for this second one
```{r}
uout1$llike
uout2$llike
```

And the allele-frequency is closer to the correct value
```{r}
rout$allele_freq
uout1$allele_freq
uout2$allele_freq
```

```{r}
sessionInfo()
```
