---
title: "Look at updot performance on overdispersion"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here, I just run through the datasets Felipe gave me.

## Read in SNPs

```{r}
library(updog)
set.seed(389)
load("../data/subset_David.Rdata")
dat <- subset_david
lastp <- 12 ## number of final rows that are parents

```

From a previous write-up, I set the sequencing error to be 0.1 percent.
```{r, cache = TRUE}
for (index in 1:10) {
  ## Run `updog`
cmat <- dat[[index]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]

p1counts <- p1dat[, 1]
p1size   <- rowSums(p1dat)
p2counts <- p2dat[, 1]
p2size   <- rowSums(p2dat)
ocounts  <- odat[, 1]
osize    <- rowSums(odat)
ploidy <- 6

uout1 <- updog(ocounts = ocounts, osize = osize, ploidy = ploidy,
              update_rho = TRUE, overdispersion = TRUE,
              update_geno = TRUE, update_outlier = TRUE,
              update_pi = TRUE)

pl <- plot_geno(ocounts = ocounts, osize = osize, ploidy = ploidy, p1counts = p1counts, p1size = p1size,
                p2counts = p2counts, p2size = p2size, col = uout1$ogeno, prob_ok = uout1$prob_ok,
                seq_error = uout1$seq_error,
                maxpostprob = apply(uout1$opostprob, 2, max))
cat("Dataset:", index, "\n")
cat("Sequencing Error:", uout1$seq_error, "\n")
cat("Overdispersion:", uout1$rho, "\n")
print(pl)

pl <- plot_geno(ocounts = ocounts, osize = osize, ploidy = ploidy, p1counts = p1counts, p1size = p1size,
                p2counts = p2counts, p2size = p2size, seq_error = uout1$seq_error)
print(pl)
}
```

I think these results look pretty good. But there is perhaps a systematic bias in a few counts (for example, datasets 5 and 7). Would it be worthwhile to model it as a random effect?

The maximum overdispersion at this ploidy was set to be `r 1/7`. None of the datasets reached this maximum.


## Session information

```{r info}
sessionInfo()
```
