---
title: "Some basic diagnostics"
author: "David Gerard"
date: "June 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract
Here, I look at the highly dispersed SNP. The log-likelihood plots don't show anything major. The average of `maxpostprob` is really low.


# Analysis

Read in Data

```{r}
library(updog)
library(tidyverse)
load("../data/subset_David.Rdata")
ploidy <- 6

cdat_list <- list()
pdat_list <- list()
for (col in 1:length(subset_david)){
  dat <- as_data_frame(subset_david[[col]])
  dat$id <- rownames(subset_david[[col]])
  names(dat) <- c("A", "a", "id")
  dat <- dat %>% mutate(ocounts = A, osize = a + A)
  pdat <- filter(dat, grepl("Beauregard", id) | grepl("Tanzania", id)) %>%
    separate("id", c("line", "id"), sep = "_") %>%
    group_by(line) %>%
    summarise(pcounts = sum(ocounts), psize = sum(osize)) %>%
    ungroup()
  cdat <- filter(dat, !(grepl("Beauregard", id) | grepl("Tanzania", id)))
  cdat_list[[col]] <- cdat
  pdat_list[[col]] <- pdat
}
```

Fit `updog`.

```{r}
cdat <- cdat_list[[3]]
pdat <- pdat_list[[3]]
uout <- updog_vanilla(ocounts = cdat$ocounts, osize = cdat$osize, 
                      ploidy = ploidy, 
                      p1counts = pdat$pcounts[1], p1size = pdat$psize[1],
                      p2counts = pdat$pcounts[2], p2size = pdat$psize[2],
                      print_val = FALSE, non_mono_max = 1)
stopifnot(dupdog(uout) == uout$llike)
plot(uout, plot_beta = FALSE, use_colorblind = TRUE)
uout$llike
```

Simulate a lot of log-likelihoods. I don't see any major discrepancy here.

```{r}
itermax <- 200
llike_vec <- rep(NA, length = itermax)
for (index in 1:itermax) {
  usim <- rupdog(uout)
  llike_vec[index] <- usim$llike  
}

llike_dat <- data_frame(llike = llike_vec)
ggplot(data = llike_dat, mapping = aes(x = llike)) +
  geom_histogram(bins = 30, fill = "white", color = "black") +
  theme_bw() +
  geom_vline(xintercept = uout$llike, lty = 2, col = 2, lwd = 2)
```

But the average `maxpostprob` is really small. We can estimate the proportion of points we expect to misclassify:
```{r}
1 - mean(uout$maxpostprob)
```

Indeed, the most sure we can possibly be about a SNP for these data is about 0.7

```{r}
  max(uout$maxpostprob)
```    

There are both many outliers and the overdispersion is large. The bias and seq-error both look ok.
```{r}
uout$out_prop
uout$bias_val
uout$seq_error
uout$od_param
plot(uout, ask = FALSE)
```


```{r}
sessionInfo()
```
