---
title: "Generate UPDOG Presentation Figures"
author: "David Gerard"
date: "February 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This file creates the figures I use for my updog presentation.


# Plot the Raw Data

```{r}
library(updog)
library(tidyverse)
load("../data/subset_David.Rdata")
ploidy <- 6

dat_list <- list()
for (col in 1:length(subset_david)){
  dat <- as_data_frame(subset_david[[col]])
  dat$id <- rownames(subset_david[[col]])
  names(dat) <- c("A", "a", "id")
  dat <- dat %>% mutate(ocounts = A, osize = a + A)
  pdat <- filter(dat, grepl("Beauregard", id) | grepl("Tanzania", id))
  cdat <- filter(dat, !(grepl("Beauregard", id) | grepl("Tanzania", id)))
  pdf(file = paste0("../output/raw_plots/geno_plot", col, ".pdf"), family = "Times", height = 4, width = 6)
  pl <- plot_geno(ocounts = cdat$ocounts, osize = cdat$osize, ploidy = ploidy, seq_error = 0)
  print(pl)
  dev.off()
  dat_list[[col]] <- cdat
  print(sort(dat$osize)[1:5])
}

```


# Argument for overdispersion

Argument for ogeno = 4.

```{r}
dat <- dat_list[[6]]
uout6 <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6, print_val = FALSE)
plot(uout6, plot_beta = FALSE)

pk_index <- 4
bindat <- filter(dat, uout6$ogeno == pk_index) %>%
  select(ocounts, osize)
pvec <- updog::get_pvec(ploidy = ploidy, bias_val = uout6$bias_val, seq_error = uout6$seq_error)

phat_naive <- sum(bindat$ocounts) / sum(bindat$osize)
phat_naive
pvec[pk_index + 1]
sd_naive <- sqrt(phat_naive * (1 - phat_naive) / sum(bindat$osize))
sd_naive ## almost zero

itermax <- 10000
simmat <- matrix(NA, nrow = itermax, ncol = nrow(bindat))
for (index in 1:itermax){
  sim_out <- rbinom(n = nrow(bindat), size = bindat$osize, prob = pvec[pk_index + 1])  
  simmat[index, ] <- sim_out
}

lower <- apply(simmat, 2, quantile, 0.025)
upper <- apply(simmat, 2, quantile, 0.975)
sim_df <- data_frame(lower = lower, upper = upper, ocounts = bindat$ocounts, osize = bindat$osize)
sim_df$mean <- pvec[pk_index + 1] * bindat$ocounts

pl <- ggplot(data = sim_df, mapping = aes(x = osize, y = ocounts)) +
  geom_point() +
  geom_line(mapping = aes(y = lower), col = "Blue") +
  geom_line(mapping = aes(y = upper), col = "Blue") +
  theme_bw() +
  xlab("Total Reads") +
  ylab("Counts A")

pdf(file = "../output/overdispersion/over4.pdf", family = "Times", height = 4, width = 6)
print(pl)
dev.off()

mean(sim_df$ocounts < lower | sim_df$ocounts > upper)

# pbinom(q = sum(sim_df$ocounts < lower | sim_df$ocounts > upper),
#        size = nrow(bindat), prob = 0.05, lower.tail = FALSE)
# pbinom(q = bindat$ocounts, size = bindat$osize, prob = pvec[pk_index], lower.tail = FALSE)

```

Same thing but using `phat_naive` with ad-hoc adjustment for estimate of phat_naive
```{r}
itermax <- 10000
simmat <- matrix(NA, nrow = itermax, ncol = nrow(bindat))
for (index in 1:itermax){
  phat_temp <- stats::rnorm(n = 1, mean = phat_naive, sd = sd_naive)
  sim_out <- rbinom(n = nrow(bindat), size = bindat$osize, prob = phat_temp)  
  simmat[index, ] <- sim_out
}

lower <- apply(simmat, 2, quantile, 0.025)
upper <- apply(simmat, 2, quantile, 0.975)
sim_df <- data_frame(lower = lower, upper = upper, ocounts = bindat$ocounts, osize = bindat$osize)
sim_df$mean <- pvec[pk_index + 1] * bindat$ocounts

pl <- ggplot(data = sim_df, mapping = aes(x = osize, y = ocounts)) +
  geom_point() +
  geom_line(mapping = aes(y = lower), col = "Blue") +
  geom_line(mapping = aes(y = upper), col = "Blue") +
  theme_bw() +
  xlab("Total Reads") +
  ylab("Counts A")

pdf(file = "../output/overdispersion/over4_alt.pdf", family = "Times", height = 4, width = 6)
print(pl)
dev.off()

mean(sim_df$ocounts < lower | sim_df$ocounts > upper)

```


Argument for ogeno = 5.

```{r}
pk_index <- 5
bindat <- filter(dat, uout6$ogeno == pk_index) %>%
  select(ocounts, osize)

itermax <- 10000
simmat <- matrix(NA, nrow = itermax, ncol = nrow(bindat))
for (index in 1:itermax){
  sim_out <- rbinom(n = nrow(bindat), size = bindat$osize, prob = pvec[pk_index + 1])  
  simmat[index, ] <- sim_out
}

lower <- apply(simmat, 2, quantile, 0.025)
upper <- apply(simmat, 2, quantile, 0.975)
sim_df <- data_frame(lower = lower, upper = upper, ocounts = bindat$ocounts, osize = bindat$osize)
sim_df$mean <- pvec[pk_index + 1] * bindat$ocounts

pl <- ggplot(data = sim_df, mapping = aes(x = osize, y = ocounts)) +
  geom_point() +
  geom_line(mapping = aes(y = lower), col = "Blue") +
  geom_line(mapping = aes(y = upper), col = "Blue") +
  theme_bw() +
  xlab("Total Reads") +
  ylab("Counts A")

pdf(file = "../output/overdispersion/over5.pdf", family = "Times", height = 4, width = 6)
print(pl)
dev.off()

mean(sim_df$ocounts < lower | sim_df$ocounts > upper)

```

Argument when ogeno = 3.

```{r}
pk_index <- 3
bindat <- filter(dat, uout6$ogeno == pk_index) %>%
  select(ocounts, osize)

itermax <- 10000
simmat <- matrix(NA, nrow = itermax, ncol = nrow(bindat))
for (index in 1:itermax){
  sim_out <- rbinom(n = nrow(bindat), size = bindat$osize, prob = pvec[pk_index + 1])  
  simmat[index, ] <- sim_out
}


lower <- apply(simmat, 2, quantile, 0.025)
upper <- apply(simmat, 2, quantile, 0.975)
sim_df <- data_frame(lower = lower, upper = upper, ocounts = bindat$ocounts, osize = bindat$osize)
sim_df$mean <- pvec[pk_index + 1] * bindat$ocounts

pl <- ggplot(data = sim_df, mapping = aes(x = osize, y = ocounts)) +
  geom_point() +
  geom_line(mapping = aes(y = lower), col = "Blue") +
  geom_line(mapping = aes(y = upper), col = "Blue") +
  theme_bw() +
  xlab("Total Reads") +
  ylab("Counts A")

pdf(file = "../output/overdispersion/over3.pdf", family = "Times", height = 4, width = 6)
print(pl)
dev.off()

mean(sim_df$ocounts < lower | sim_df$ocounts > upper)

```


Plot a lot of betas
```{r}
rhoseq <- c(0.005, 0.01, 0.05, 0.1, 0.1666, 0.2)
mu <- 0.2
dat <- data_frame(x = as.numeric(c()), y = as.numeric(c()))
for (index in 1:length(rhoseq)) {
  rho <- rhoseq[index]  
  alpha <- mu * (1 - rho)/rho
  beta <- (1 - mu) * (1 - rho)/rho
  x <- seq(0.0, 1, length = 100)
  y <- dbeta(x, alpha, beta)
  tempdat <- data_frame(x = x, y = y)
  tempdat$rho <- rho
  dat <- bind_rows(dat, tempdat)
}
dat$rho <- as.factor(paste0("p = 0.2, s = ", dat$rho))


pl <- ggplot(data = dat, mapping = aes(x = x, y = y)) +
  facet_wrap(~rho) +
  geom_line() +
  theme_bw() +
  ylab("Density") +
  xlab("q") +
  theme(strip.background = element_rect(fill = "white")) +
  geom_vline(xintercept = mu, lty = 2, col = "grey")
pdf(file = "../output/overdispersion/beta_plots.pdf", family = "Times", height = 4, width = 6)
print(pl)
dev.off()

```

# Argument for outliers

```{r}
dat <- dat_list[[10]]
uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6)

uout$p1geno
uout$p2geno

dat$osize[dat$ocounts <= 5]
dat$ocounts[dat$ocounts <= 5]

possible_probs <- updog::get_pvec(ploidy = ploidy, bias_val = uout$bias_val, 
                                  seq_error = uout$seq_error)

## Probabilities of points under fit model
library(VGAM)
VGAM::pbetabinom(q = 0, size = 116, prob = possible_probs[5], rho = uout$od_param)
VGAM::pbetabinom(q = 2, size = 116, prob = possible_probs[5], rho = uout$od_param)
VGAM::pbetabinom(q = 1, size = 84, prob = possible_probs[5], rho = uout$od_param)

## Probabilities of points under naive binomial model
pbinom(0, 116, 2/3)
pbinom(2, 116, 2/3)
pbinom(1, 84, 2/3)

```

# Fit updog

Fit all updog
```{r}
for(col in 1:length(dat_list)) {
  dat <- dat_list[[col]]
  uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6, print_val = FALSE,
                        seq_error_sd = Inf, bias_val_sd = Inf)

  pdf(file = paste0("../output/updog_plots/uout", col, ".pdf"), family = "Times", height = 5, width = 5)
  plot(uout, plot_beta = FALSE)
  dev.off()
  
  pl <- plot_geno(ocounts = dat$ocounts, osize = dat$osize, ploidy = ploidy, seq_error = 0)
  pdf(file = paste0("../output/updog_plots/rout", col, ".pdf"), family = "Times", height = 5, width = 5)
  print(pl)
  dev.off()
}
```

```{r}
## no overdispersion
col <- 6
dat <- dat_list[[col]]
uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6,
                      update_od_param = FALSE,
                      od_param = 10^-7)

pdf(file = paste0("../output/updog_plots/uout_alt", col, ".pdf"), family = "Times", height = 5, width = 5)
plot(uout, plot_beta = FALSE)
dev.off()
```


How likely are some aaaAAA child genotypes when AAAAAA and aaaAAA?
```{r}
qout <- get_q_array(6)
qout[4, 7, 4] * nrow(dat)
qout[5, 6, 4] * nrow(dat)
```


Look at how likely each obs is an outlier in dat_list[[3]]
```{r}
dat <- dat_list[[3]]
uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6)
1 - uout$out_prop
plot_beta_dist_gg(mu = uout$out_mean, rho = uout$out_disp)


## Plot the overdispersed beta
mu <- uout$out_mean
rho <- uout$out_disp
alpha <- mu * (1 - rho)/rho
beta <- (1 - mu) * (1 - rho)/rho
x <- seq(0.0, 1, length = 100)
y <- dbeta(x, alpha, beta)
tempdat <- data_frame(x = x, y = y)
pl <- ggplot(data = tempdat, mapping = aes(x = x, y = y)) +
  geom_line() +
  theme_bw() +
  ylab("Density") +
  xlab("q") +
  geom_vline(xintercept = mu, lty = 2, col = "grey") +
  ylim(0, 1.4)
```


# Now fit updog while updating outlier distribution

```{r}
for(col in 1:length(dat_list)) {
  dat <- dat_list[[col]]
  uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6, update_outdisp = TRUE,
                        update_outmean = TRUE, bound = FALSE, non_mono_max = 200, print_val = FALSE)

  pdf(file = paste0("../output/updog_plots/over_uout", col, ".pdf"), family = "Times", height = 5, width = 5)
  plot(uout, plot_beta = FALSE)
  dev.off()
}
```



```{r}
## Seven works poorly because of updating outlier
dat <- dat_list[[7]]
uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6,
                      update_outdisp = TRUE,
                      update_outmean = TRUE, bound = FALSE, 
                      print_val = FALSE)
pdf(file = paste0("../output/updog_plots/over_uout", 7, "_beta.pdf"), family = "Times", height = 5, width = 5)
alpha <- uout$out_mean * (1 - uout$out_disp) / uout$out_disp
beta <- (1 - uout$out_mean) * (1 - uout$out_disp) / uout$out_disp
plot_beta_dist_gg(alpha = alpha, beta = beta)
dev.off()
uout$out_prop
```

Does the fit for 7 now more-closely match the segregation probabilities?
```{r}
xtab <- matrix(NA, nrow = 2, ncol = 2)
ogeno_vec <- uout$ogeno
xtab[1, ] <- table(ogeno_vec) / sum(!is.na(ogeno_vec))
ogeno_vec[is.na(ogeno_vec)] <- 5
xtab[2, ] <- table(ogeno_vec) / length(ogeno_vec)
xtable::xtable(xtab)
get_q_array(6)[7, 6, ]

colMeans(uout$postmat)
```

```{r}
## Ten works fine with bound
dat <- dat_list[[10]]
uout <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6,
                      update_outdisp = TRUE,
                      update_outmean = TRUE, bound_bias = FALSE, 
                      print_val = FALSE)
plot(uout, ask = FALSE)

uout2 <- updog_vanilla(ocounts = dat$ocounts, osize = dat$osize, ploidy = 6, update_outdisp = TRUE,
                      update_outmean = TRUE, bound = TRUE, non_mono_max = 200, print_val = FALSE,
                      bias_val = uout$bias_val, od_param = uout$od_param, 
                      seq_error = uout$seq_error, out_prop = uout$out_prop)
plot(uout, ask = FALSE)
```

```{r}
sessionInfo()
```
