---
title: "Get Seq Error Prior"
author: "David Gerard"
date: "June 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I read in the sequencing error that Blischak used and inflate it a little to derive my default prior for `updog`.

# Analysis

```{r}
seq_error1 <- read.table("~/Code/polyploid-genotyping/data/c20-pendula-error.txt")[, 1]
seq_error2 <- read.table("~/Code/polyploid-genotyping/data/c20-pubescens-error.txt")[, 1]

hist(seq_error1)
hist(seq_error2)
summary(seq_error1)
summary(seq_error2)
```

Use method of moments to get prior values on mean and overdispersion parameters assuming a beta.

```{r}
mu = mean(seq_error1)
tau = var(seq_error1) / (mu * (1 - mu))
mu
tau

mu = mean(seq_error2)
tau = var(seq_error2) / (mu * (1 - mu))
mu
tau
```

It seems taht setting $\mu = \tau = 0.005$ would be a reasonable prior on the sequencing error rate. We can inflate $\tau$ a little bit to 0.01 or even 0.02 or 0.03 or 0.1 to be safe.

```{r}
updog::plot_beta_dist(mu = 0.005, rho = 0.01)
updog::plot_beta_dist(mu = 0.005, rho = 0.02)
updog::plot_beta_dist(mu = 0.005, rho = 0.03)
updog::plot_beta_dist(mu = 0.005, rho = 0.04)
updog::plot_beta_dist(mu = 0.005, rho = 0.05)
updog::plot_beta_dist(mu = 0.005, rho = 0.1)
```

See if the maximum sequencing error observed in the Blischak data is consistant with this prior.

```{r}
max1 <- max(seq_error1)
max2 <- max(seq_error2)
max1
max2

mu <- 0.005
tau <- 0.02

alpha <- mu * (1 - tau) / tau
beta  <- (1 - mu) * (1 - tau) / tau
pupper <- pbeta(q = max1, shape1 = alpha, shape2 = beta, lower.tail = FALSE)
pupper
```

Having a `r pupper` upper tail probability is still pretty large considering that we have `r length(seq_error1)` points.


```{r}
sessionInfo()
```
