---
title: "Look at updated estimate of sequencing error"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here, I demonstrate some new methods for the output of `updog`.

## Demonstration

```{r}
library(updog)
set.seed(483)
load("../data/subset_David.Rdata")
dat <- subset_david
lastp <- 12 ## number of final rows that are parents

## Run `updog`
cmat <- dat[[2]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]

p1counts <- p1dat[, 1]
p1size   <- rowSums(p1dat)
p2counts <- p2dat[, 1]
p2size   <- rowSums(p2dat)
ocounts  <- odat[, 1]
osize    <- rowSums(odat)
ploidy <- 6
```

Now I fit updog.

```{r}
uout <- updog(ocounts = ocounts, osize = osize, ploidy = ploidy, update_geno = TRUE)
```

After fitting updog, you can now just type `plot` on the output and get good genotype plots:
```{r}
plot(uout)
```

The second plot above is the estimated beta distribution for the outlier model. The third plot is for the overdisperion model for each possible genotype. You can ignore these with `plot_beta = FALSE`.

If you prefer "classic" base R graphics, you can set `gg = FALSE`.

```{r}
plot(uout, gg = FALSE, plot_beta = FALSE)
```

The `summary` method now works with the `updog` output.
```{r}
summary(uout)
```

```{r info}
sessionInfo()
```
