---
title: "First SNP Analaysis"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This is a template for writing reports with R Markdown.

## Read in SNPs and look at data

```{r}
library(updog)
set.seed(3452)
load("../data/subset_David.Rdata")
dat <- subset_david
lastp <- 12 ## number of final rows that are parents

for(index in 1:length(dat)) {
  cmat  <- dat[[index]]
  p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
  p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
  odat  <- cmat[1:(nrow(cmat) - lastp), ]
  plot_geno(ocounts = odat[, 1], osize = odat[, 2] + odat[, 1], ploidy = 6,
            p1counts = p1dat[, 1], p1size = p1dat[, 1] + p1dat[, 2],
            p2counts = p2dat[, 1], p2size = p2dat[, 1] + p2dat[, 2])
}


```

In the last three datasets, there appear to be three two observations that do not align with the rest of the data. 
This is curious, but they are not the same observation.
```{r}
cmat <- dat[[8]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
cmat <- dat[[9]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
cmat <- dat[[10]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
```

## Run `updog`

We now run updog both with and without parental genotypes.
```{r}
cmat <- dat[[10]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]
uout <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
              p1counts = p1dat[, 1], p1size = rowSums(p1dat),
              p2counts = p2dat[, 1], p2size = rowSums(p2dat))

uout2 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
               p1counts = p1dat[, 1], p1size = rowSums(p1dat),
               p2counts = p2dat[, 1], p2size = rowSums(p2dat),
               do_mcmc = TRUE)

uout3 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
               do_mcmc = TRUE)
```

The parental genotypes are chosen with near certainty
```{r}
aprobs <- cmat[1:(nrow(cmat) - 12), 1]/ rowSums(cmat[1:(nrow(cmat) - 12),])
corder <- order(aprobs)
uout$p1postprob
uout$p2postprob
uout2$m_p1postprob
uout2$m_p2postprob

aprobs[corder[1:4]]
uout$opostprob[, corder[1:4]]
uout2$m_opostprob[, corder[1:4]]
```

Plot output
```{r}
labelvec <- apply(uout$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

labelvec <- apply(uout2$m_opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

labelvec <- apply(uout3$m_opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                col = labelvec)

```

It seems that the MCMC gives me what Felipe wants when I don't use the parental data, this is because the few extreme offspring points draw one of the parents to have an "a" when the sequence data says that it shouldn't:
```{r}
cbind(uout3$m_p1postprob,
      uout2$m_p1postprob,
      round(uout$p1postprob, digits = 2))

cbind(uout3$m_p2postprob,
      uout2$m_p2postprob,
      round(uout$p2postprob, digits = 2))
```


I am getting very different results from the MCMC for the parental genotypes depending on the seed I use in the case when the parental genotypes are not provided. This might be because of the identifiability issue in the parents labeling, but I don't think so because we see probabilities near 1:
```{r}
set.seed(195)
getmc1 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
                do_mcmc = TRUE)
set.seed(196)
getmc2 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
                do_mcmc = TRUE)

cbind(getmc1$m_p1postprob, getmc2$m_p1postprob)
cbind(getmc1$m_p2postprob, getmc2$m_p2postprob)

plot(getmc1$m_opostprob, getmc2$m_opostprob)
```

## Take aways:

1. There's a bug in my MCMC and I need to fix it.
2. We should probably model a fudge factor.






## Session information

```{r info}
sessionInfo()
```
