---
title: "First SNP Analaysis"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

## Read in SNPs and look at data

```{r}
library(updog)
set.seed(3452)
load("../data/subset_David.Rdata")
dat <- subset_david
lastp <- 12 ## number of final rows that are parents

for(index in 1:length(dat)) {
  cmat  <- dat[[index]]
  p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
  p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
  odat  <- cmat[1:(nrow(cmat) - lastp), ]
  plot_geno(ocounts = odat[, 1], osize = odat[, 2] + odat[, 1], ploidy = 6,
            p1counts = p1dat[, 1], p1size = p1dat[, 1] + p1dat[, 2],
            p2counts = p2dat[, 1], p2size = p2dat[, 1] + p2dat[, 2])
}


```

In the last three datasets, there appear to be three two observations that do not align with the rest of the data. 
This is curious, but they are not the same observation.
```{r}
cmat <- dat[[8]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
cmat <- dat[[9]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
cmat <- dat[[10]]
sort(cmat[, 1] / (cmat[, 1] + cmat[, 2]))[1:4]
```

## Run `updog`

We now run updog both with and without parental genotypes.
```{r}
cmat <- dat[[10]]
p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
odat  <- cmat[1:(nrow(cmat) - lastp), ]
uout <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
              p1counts = p1dat[, 1], p1size = rowSums(p1dat),
              p2counts = p2dat[, 1], p2size = rowSums(p2dat))

uout2 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6,
               p1counts = p1dat[, 1], p1size = rowSums(p1dat),
               p2counts = p2dat[, 1], p2size = rowSums(p2dat))

uout3 <- updog(ocounts = odat[, 1], osize = rowSums(odat), ploidy = 6)
```

The parental genotypes are chosen with near certainty
```{r}
aprobs <- cmat[1:(nrow(cmat) - 12), 1]/ rowSums(cmat[1:(nrow(cmat) - 12),])
corder <- order(aprobs)
uout$p1postprob
uout$p2postprob
uout2$m_p1postprob
uout2$m_p2postprob

aprobs[corder[1:4]]
uout$opostprob[, corder[1:4]]
uout2$m_opostprob[, corder[1:4]]
```

Plot output
```{r}
labelvec <- apply(uout$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                ogeno = labelvec)
print(pl)

labelvec <- apply(uout2$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                ogeno = labelvec)
print(pl)

labelvec <- apply(uout3$opostprob, 2, which.max)
pl <- plot_geno(ocounts = odat[, 1], osize = rowSums(odat),
                p1counts = p1dat[, 1], p1size = rowSums(p1dat),
                p2counts = p2dat[, 1], p2size = rowSums(p2dat),
                ploidy = 6,
                ogeno = labelvec)
print(pl)

```





## Session information

```{r info}
sessionInfo()
```
