---
title: "Try ASH on updog fits"
author: "David Gerard"
date: "June 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I just play around with fitting ASH onthe fits from `updog`.

```{r}
library(updog)
library(tidyverse)
numfiles <- length(list.files("~/Code/reproduce_genotyping/Output/updog_fits/"))
parmat <- matrix(NA, nrow = numfiles, ncol = 9)
colnames(parmat) <- c("bias", "seq_error", "od", "s", "ell", "r", "se_s", "se_ell", "se_r")
for (index in 1:numfiles) {
  uout <- readRDS(paste0("~/Code/reproduce_genotyping/Output/updog_fits/uout", index, ".RDS"))
  parmat[index, 1]   <- uout$bias_val
  parmat[index, 2]   <- uout$seq_error
  parmat[index, 3]   <- uout$od_param
  parmat[index, 4]   <- log(uout$bias_val)
  parmat[index, 5]   <- log(uout$seq_error / (1 - uout$seq_error))
  parmat[index, 6]   <- log((1 - uout$od_param) / uout$od_param)
  parmat[index, 7:9] <- sqrt(-1 * diag(solve(uout$hessian)))
}
pardat <- as_data_frame(parmat)
```

There are a couple points with a relatively large standard error in s and r.
```{r}
qplot(pardat$se_s, bins = 30)
qplot(pardat$se_ell, bins = 30)
qplot(pardat$se_r, bins = 30)
```

## ASH on Bias
```{r}
library(ashr)
aout <- ash(betahat = pardat$s, sebetahat = pardat$se_s)
```

There are a couple points that really move toward 1:
```{r}
bias_est <- exp(get_pm(aout))
qplot(pardat$bias, bias_est) +
  geom_abline(color = "red")
```

## ASH on Sequencing Error Rate
```{r}
aout <- ash(betahat = pardat$ell, sebetahat = pardat$se_ell, mode = "estimate", mixcompdist = "halfuniform")
```

There are a couple points that really move toward 1:
```{r}
seq_est <- exp(get_pm(aout)) / (1 + exp(get_pm(aout)))
qplot(pardat$seq_error, seq_est) +
  geom_abline(color = "red")
```

## ASH on OD param
```{r}
aout <- ash(betahat = pardat$r, sebetahat = pardat$se_r, mode = "estimate", mixcompdist = "halfuniform")
```

```{r}
od_est <- exp(-1 * get_pm(aout)) / (1 + exp(-1 * get_pm(aout)))
qplot(pardat$od, od_est) +
  geom_abline(color = "red")
```

# Refit updog with new fits
We will compare fits where the differences in OD are very large

```{r}
index <- which(pardat$od > 0.007 & od_est < 0.005)
uout <- readRDS(paste0("~/Code/reproduce_genotyping/Output/updog_fits/uout", index, ".RDS"))
unew <- updog_vanilla(ocounts = uout$input$ocounts, osize = uout$input$osize, ploidy = uout$input$ploidy, 
                      p1counts = uout$input$p1counts, p1size = uout$input$p1size, od_param = od_est[index],
                      bias_val = bias_est[index], seq_error = seq_est[index], update_od_param = FALSE,
                      update_bias_val = FALSE, update_seq_error = FALSE)
uout$od_param
unew$od_param
```

Minor difference in genotyping
```{r}
sum(uout$ogeno != unew$ogeno)
```

It's just a low-count point that changes genotype.
```{r}
plot(uout, plot_beta = FALSE)
plot(unew, plot_beta = FALSE)
```


```{r}
sessionInfo()
```
