---
title: "First Fits with Outlier Model"
author: "First Last"
date: "`r Sys.Date()`"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

Here, I try out vanilla updog with the outlier model where the outlier model is Unif(0,1). This works well if `out_prop` is initially set to `0.001` but can work very poorly if `out_prop` is initialized too large. I need to worry about the large likelihood values I get for SNP 10 when I initialize at `out_prop = 0.01`.

## Read in Data

```{r}
library(updog)
set.seed(483)
load("../data/subset_David.Rdata")
ddat <- subset_david
lastp <- 12 ## number of final rows that are parents
```


```{r, cache=TRUE}
for (index in 1:length(ddat)) {
  cmat <- ddat[[index]]
  p1dat <- cmat[(nrow(cmat) - lastp + 1):(nrow(cmat) - lastp / 2), ]
  p2dat <- cmat[(nrow(cmat) - lastp / 2 + 1):nrow(cmat), ]
  odat  <- cmat[1:(nrow(cmat) - lastp), ]

  p1counts <- p1dat[, 1]
  p1size   <- rowSums(p1dat)
  p2counts <- p2dat[, 1]
  p2size   <- rowSums(p2dat)
  ocounts  <- odat[, 1]
  osize    <- rowSums(odat)
  ploidy <- 6

  uout <- updog_vanilla(ocounts = ocounts, osize = osize, ploidy = ploidy, min_disp = 0,
                        print_val = FALSE, commit_num = 4)
  print(plot_geno(ocounts = ocounts, osize = osize, ploidy = ploidy, seq_error = 0))
  plot(uout, ask = FALSE, plot_beta = TRUE)
  
}
```

```{r info}
sessionInfo()
```
